---
title       : Task Stressors
subtitle    : OHP
author      : Shane Curtis
job         : 
framework   :	io2012
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : [bootstrap, quiz]            # {mathjax, quiz, bootstrap}
mode        : selfcontained	# {standalone, draft, selfcontained}
ext_widgets	: {rCharts: [libraries/nvd3]}
---

## Overview

> - What are task stressors?
> - How are they different from/similar to other stressors
> - How do we study them
> - Research Framework
> - Research findings

---

## Examples

> - Weekly work schedule
> - Over- and underutilization of skills
> - Variance in workload
> - Pace of work
> - Responsibility (for people and things)
> - Travel as part of the job
> - Job Characteristics

---

## Conceptual Overlap

> - Organizational constraints
> - Role conflict
> - Work-family conflict

*** pnotes

- prevent workers from maintaining their typical level of performance
- Multiple roles requires more time and effort

---

## Shiftwork

> - Adversely affects sleep
> - Promotes fatigue
> - Associated with the occurence of accidents and injuries
> - Psychological distress
> - Cardiovascular disorders

---

## Workload

> - Quantitative vs. Qualitative
> - Mental vs. Physical
> - Objective vs. Perceived
> - Naturally-occuring vs. Experimentally-manipulated

---

## Quantitative vs. Qualitative

> - Amount
> 	- Hours worked

> - Difficulty
> 	- Physical
> 	- Mental
> 	- Emotional

---

## Mental vs Physical

> - Different causes and outcomes

---

## Objective vs Perceived

> - Verifiable amount or difficulty of work
> 	- "How many hours do you work during a typical week?"
> - Relative amount or difficulty of work
> 	- Subjective individual standard
> 		- Ability
> 		- Time
> 		- Preference
> 	- "The amount of work I am expected to do is too great"

---

## Naturally-occuring vs Experimentally-manipulated

> - Characteristics of one's job as a whole
> - Characteristics of the particular work task

*** pnotes

Generalizability
unable to examine causal effects

---

## How are they Studied

> - Often studied alongside other stressors
> - Makes it difficult to interpret research findings

---

## Research Framework

<iframe src ='assets/img/research.png' width=250px height=600px></iframe>

---

## Potential Causes of Excessive Workload

> - Work Environment
> - Individual Differences

---

## Work Environment

> - Leader Behaviors
>		- assign too much/little work
> - Culture
> 	- Organizational level
> 	- Occupational level
> - Varying Patterns of Work Demands
>		- creates unpredictability

*** pnotes

Tax accountant
Teacher

---

## Individual Differences

> - Ability
> - Time Management Skills
> - Negative Affectivity
> - Conscientiousness

---

## Ability

> - Increase quantitative/qualitative workload
> - Decrease percieved workload
> - Increase objective workload

---

## Time Management Skills

> - Can cause work to "pile up"
> - Relavance varies by job

---

## Negative Affectivity

> - Positively related to perceived workload
> - Unrelated to objective workload

---

## Conscientiousness

> - Volunteer to do more work
> - Not necessarily bad

--- .title-only 

# Outcomes

---

## Effects on Employee Well Being

> - Psychological Strains
> - Physical Strains

---

## Psychological Strains

> - Anxiety (.40)
> - Depression (.21)
> - Frustration (.46)

---

## Physical Strains

> - Physical Health Symptoms (.27)

---

## Effects on Employee Behavior

> - In-Role Job Performance
> - OCBs (.31)
> - CWBs (.21)
> - Withdrawal

---

## Job Attitudes

> - Engagement
> - Job Satisfaction

---

```{r, echo=FALSE, results='hide', comment=NA, message=FALSE}
require(MASS)
rmat <- matrix(c(1.0, .40, .21, .46, .27, .31, .21, .40, 1.0, .30, .25, .40, .02, .10, .21, .30, 1.0, .10, .10, -.20, .20, .46, .25, .10, 1.0, .15, -.15, .12, .27, .40, .10, .15, 1.0, .03, .10, .31, .02, -.20, -.15, .03, 1.0, -.30, .21, .10, .20, .12, .10, -.30, 1.0), nrow = 7, byrow = T)
mu <- c(0,0,0,0,0,0,0)
mat <- data.frame(mvrnorm(200, Sigma = rmat, mu = mu, empirical = TRUE))
names(mat) <- c("Workload", "Anxiety", "Depression", "Frustration", "PHS", "OCB", "CWB")
```

```{r inter-plot, echo=FALSE, results='asis', comment=NA, message=FALSE}
require(rCharts)
r1 <- rPlot(Anxiety ~ Workload, data = mat, type='point')
r1$addControls("y", "Anxiety", names(mat))
r1
```

---

## Two Qualifactions

> - Cross-sectional non-experimental
> - Assumes a linear relationship

---

## Combating Organizational Problems

> - Selection
> - Training
> - Job Redesign

---

## Selection

> - Misfit between the KSAs of individual workers and the demands of the job
> 	- High qualitative workload
> 	- inadequate staffing

---

## Training

> - Help employees develop the KSAs required by the job
> - Train managers to eliminate ineffective behaviors

---

## Job Redesign

> - Redistribute work among employees



